Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-05-28 cvx-flats-neurips21/cvx-flats.tex
Spot check after submission

\documentclass[11pt]{article} % Anonymized submission
% \renewcommand{\includegraphics}[1][1]{}

@@

% examples of problems where we care a lot about prediction dimension (that happen to be finite discrete)
Despite the ubiquity of surrogate losses, we lack general frameworks to design and analyze consistent surrogates.
This state of affairs is especially dire when one seeks low \emph{prediction dimension}, the dimension of the surrogate prediction domain.
For example, in multiclass classification with $n$ labels, the prediction domain might be $\reals^n$.
In many type (I) settings, such as structured prediction and extreme classification, the prediction dimension of any convex and consistent surrogate often becomes intractably large, forcing one to sacrifice consistency for computational efficiency.
To understand whether this sacrifice is necessary, recent work developed tools like the feasible subspace dimension to lower bound the prediction dimension of any consistent convex surrogate~\citep{ramaswamy2016convex}.
Challenges of type (II) include estimating risk measures such as conditional value at risk (CVaR), with applications in financial regulation, robust engineering design, and algorithmic fairness.
\jessiet{cites?}
\raft{Yeah... let's figure this out after submission, since they are at least in \S~\ref{sec:risk-measures}.}
Risk measures are not elicitable, meaning they cannot be specified via a target loss, and thus we seek a surrogate loss of low (or at least finite) prediction dimension.
Recent work~\citep{fissler2016higher,frongillo2015elicitation,frongillo2020elicitation} gives prediction dimension bounds for some of these risk measures, but without the requirement that the surrogate be convex; bounds for convex surrogates are left as a major open question.

@@

\begin{restatable}{theorem}{convexflats}  \label{thm:convex-flats-inf-dim}
  Let $\Gamma:\P \toto \reals^d$ be (directly) elicited by $L \in \Lcvx_d$ for some $d\in\mathbb{N}$.
  Let $\Y$ be either a finite set, or $\Y = \reals$, in which case we assume each $p\in\P$ admits a Lebesgue density supported on the same set for all $p\in\P$.%
  \footnote{This assumption is largely for technical convenience, to ensure that $\V_{u,p}$ does not depend on $p$.
    Any such assumption would suffice, and we suspect even that condition can be relaxed.}
  For all $u\in\range\Gamma$ and $p\in\Gamma_u$, there is some $d$-flat $F$ such that $p \in F \subseteq \Gamma_u$.
 % For all $u\in\range\Gamma$ and $p\in\Gamma_u$, there is some $V_{u,p}:\Y\to\reals^d$ such that $p \in \zeros{V_{u,p}} \subseteq \Gamma_u$.
\btw{Raf: let's ditch $V_{u,p}$ in the statement, and instead say ``there is some $d$-flat $F$ such that $p \in F \subseteq \Gamma_u$'', since (a) that's the way we write it everywhere else, (b) we justify the story that our main tool is about $d$-flats, and (c) it becomes clearer why the two Corollaries could follow.  Am I missing a reason to keep $V_{u,p}$? Jessie: I like this change; kept the old version in a comment, but no reason I can think of to keep $V_{u,p}$}
\end{restatable}
\begin{proof}[Proof (finite case)]
  We will prove the result for the case of finite $\Y$, and defer the $\Y=\reals$ case to \S~\ref{app:implications-of-bounds}.
  As $L$ is convex and elicits $\Gamma$, we have $u \in \Gamma(p) \iff \vec 0 \in \partial \exploss{L}{u}{p}$. 
  With $\Y$ finite, this is additionally equivalent to $\vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the Minkowski sum~\citep[Theorem 4.1.1]{hiriart2012fundamentals}.\footnote{$\partial$ represents the subdifferential $\partial f(x) = \{z : f(x') - f(x) \geq \inprod{z}{x'-x}\; \forall x' \}$.}
  Expanding, we have $\oplus_y p_y \partial L(u,y) = \{ \sum_{y\in\Y} p_y x_y \mid x_y \in \partial L(u,y) \; \forall y\in\Y\}$, and thus there is a $W$ such that $W p = \sum_y p_y x_y = \vec 0$ where $W = [x_1, \ldots, x_n] \in \reals^{d\times n}$; cf.~\cite[$\mathbf{A}^m$ in Theorem 16]{ramaswamy2016convex}.
  Let $V_{u,p} : \Y \to \reals^d, y \mapsto W_y$ be the function encoding the columns of $W$.
  Observe that $\E_p V_{u,p} = \vec 0$.				 
  We take the flat $F := \zeros{V_{u,p}}$, and have $p \in F$ by construction.
  To see $F \subseteq \Gamma_u$, from the chain of equivalences above, we have for any $q\in\P$ that $q \in \zeros{V_{u,p}} \implies \vec 0 \in \partial \E_q L(u,Y) \implies u \in \Gamma(q) \implies q \in \Gamma_u$.
\end{proof}

@@

\paragraph{Q4: Variance}
Consider the task of estimating the variance $\Var(p) = \E_p[Y^2] - \E_p[Y]^2$.
The variance is not (directly) elicitable as its level sets are not convex~\cite{osband1985providing,lambert2008eliciting}, meaning this task falls under Quadrant 4.
Interestingly, the fact that the variance is not elicitable does not yield a lower bound on elicitation complexity of 2, as it does not rule out the variance being a link of a real-valued convex-elicitable property; cf.~\citet[Remark 1]{frongillo2020elicitation}.
In \S~\ref{app:variance}, we show $\eliccvx(\Var)=2$, meaning the lowest dimension of a convex loss to estimate conditional variance is 2. %~\citep{osband1985providing,lambert2018elicitation}.
This lower bound will follow from Theorem~\ref{thm:bayes-risk-lower-bound} in \S~\ref{sec:contin-consis} using the fact that variance is the Bayes risk of squared loss. % $L(r,y) = (r-y)^2$, which elicits the mean $\Gamma(p) = \E_p Y$.
While perhaps intuitively obvious, even this simple result is novel.

@@

Using our $d$-flats framework, we resolve both open questions with new lower bounds in both settings.
% The central observation is that the level sets of $d$-dimensional convex-elicitable properties, if not $d$-flats themselves, are
Our framework clarifies the relationship between $d$-identifiable properties and properties elicited by $d$-dimensional convex losses: the level sets of the former are $d$-flats by definition, while the level sets of the latter are \emph{unions} of $d$-flats by Theorem~\ref{thm:convex-flats-inf-dim}.
A careful examination of the arguments of Frongillo and Kash~\citep{frongillo2020elicitation,frongillo2015elicitation} and~\citet{dearborn2020indirect} reveals that they largely rely on the containment of $d$-flats in level sets, rather than the full structure of identifiable properties.
As such, although quite subtle in the case of risk measures, the general structure of these previous proofs go through for convex-elicitable properties: since no $d$-flat could be contained in a particular level set, no union of $d$-flats could be either.
Our lower bounds therefore match both of these papers, though we conjecture that our convex consistency bounds could be tightened in some cases.

@@

The problem of estimating a risk or uncertainty measure of $Y$ is of central importance in financial regulation~\cite{acerbi2014back-testing,fissler2016expected,campbell2005review} and robust engineering design~\cite{shahbaz2016surrogate-based,beyer2007robust,rockafellar2010buffered}.
% , and algorithmic fairness~\citep{williamson2019fairness}
Risk measures include the upper confidence bound $\E[Y] + \lambda \sqrt{\Var[Y]}$, or the conditional value at risk (CVaR) defined below in eq.~\eqref{eq:elic-complex-es-3}, in either conditional or unconditional contexts.
Uncertainty measures include the variance, entropy, or norm of the distribution of $Y$.
Risk and uncertainty measures are typically not elicitable, so this problem falls under Quadrant 4.
Frongillo and Kash~\citep{frongillo2020elicitation,frongillo2015elicitation} give prediction dimension lower bounds for a broad class of risk and uncertainty measures, namely Bayes risks.
As stated above, these bounds are with respect to identifiable properties, and bounds for convex surrogates are left as a major open question.

@@

To illustrate the theorem, we briefly apply it to one of the most prominent financial risk measures, the conditional value at risk (CVaR).
Several other applications from Frongillo and Kash~\citep{frongillo2020elicitation,frongillo2015elicitation}, such as other risk measures, entropy, and norms, follow similarly.
The authors observe that CVaR can be expressed as a Bayes risk; for $0 < \alpha < 1$, we may define
\begin{align}
  \CVaR_\alpha(p)
  &= \inf_{r\in\reals} \E_p\left\{ \tfrac 1 \alpha(r-Y)\ones_{r\geq Y}- r\right\}~,
  \label{eq:elic-complex-es-3}
\end{align}
which is the Bayes risk of the transformed pinball loss $L_\alpha(r,y) = \tfrac 1 \alpha(r-y)\ones_{r\geq y}-r$.
In turn, $L_\alpha$ elicits the $\alpha$-quantile, the quantity $q_\alpha(p)$ such that $\Pr_p[Y \geq q_\alpha(p)] = \alpha$.
Following \citet{frongillo2020elicitation}, we will restrict to the set $\P_q$ of probability measures over $\reals$ with connected support and whose CDFs are strictly increasing on their support, so that $q_\alpha$ is single-valued.
Under mild assumptions, we find that there is no consistent real-valued convex surrogate for $\CVaR_\alpha$.

@@

Several important questions remain open.
Particularly for the discrete settings, we would like to know whether one can lift the restriction that surrogates always achieve a minimum; we conjecture positively (see \S~\ref{app:minnable-assumption}).
% For continuous settings, it would be especially interesting to extend to the case $\Y=\reals$, which would require a more careful analysis in Theorem~\ref{thm:bayes-risk-lower-bound}.\jessie{Flagging this bc we take care of it I hope}
The observation that our bounds are as tight as calibration-based bounds, yet we use the weaker condition of indirect elicitation, motivates the study of how much weaker indirect elicitation is than calibration.
More broadly, we would like to characterize $\conscvx$ and $\eliccvx$ and develop a general framework for constructing surrogates achieving the best possible prediction dimension.

% Moreover, the practical reason why consistency is desired is to ensure the guarantee of empirical risk minimization (ERM) rates; however, the relationship between ERM rates and property elicitation has not been studied.
% \jessie{Not sure if there will be a citation for the rates paper...?}
%One might be able to further tighten these bounds by property elicitation by studying monotonicity and adjacency of level sets.
%In discrete predictions, these bounds might also be tightened if the equivalence of convex consistency dimension and embedding dimension of~\citet{finocchiaro2020embedding} is shown; the current embedding dimension bounds are not tight either, but additional structure is imposed by considering the embedding framework.

@@

\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\section{Consistency Implies Indirect Elicitation}\label{sec:consis-implies-indir}

@@

\section{Implications of Convex Indirect Elicitation Bounds}\label{app:implications-of-bounds}

@@

\section{Definitions of Calibration}\label{app:calibration}

When given a discrete target loss, such as for classification-like problems, direct empirical risk minimization is typically NP-hard, forcing one to find a more tractable surrogate.
To ensure consistency, the literature has embraced the notion of \emph{calibration} from~\citet[Chapter 3]{steinwart2008support}, which aligns with the definition in~\citet{tewari2007consistency} for multiclass classification, and its generalizations to arbitrary discrete target losses~\citep{agarwal2015consistent,ramaswamy2016convex}.
Calibration is more tractable and weaker than consistency, yet the two are equivalent under suitable assumptions~\citep{tewari2007consistency,ramaswamy2016convex}, notably in Quadrant 1.
Intuitively, calibration says one cannot achieve the optimal surrogate loss while linking to a suboptimal target prediction.

@@

\section{Quadrant 1: Previous Lower Bounds and Comparisons}\label{sec:finite-calib}

@@

Throughout this section, we will assume $\P$ is convex.
See \citet[\S E.5]{frongillo2020elicitation} for a discussion of how to relax this assumption.

@@

\section{Omitted Discussion and Examples}\label{app:omitted}

>>>>

\documentclass{article} % Anonymized submission
% \renewcommand{\includegraphics}[1][1]{}

@@

% examples of problems where we care a lot about prediction dimension (that happen to be finite discrete)
Despite the ubiquity of surrogate losses, we lack general frameworks to design and analyze consistent surrogates.
This state of affairs is especially dire when one seeks low \emph{prediction dimension}, the dimension of the surrogate prediction domain.
For example, in multiclass classification with $n$ labels, the prediction domain might be $\reals^n$.
In many type (I) settings, such as structured prediction and extreme classification, the prediction dimension of any convex and consistent surrogate often becomes intractably large, forcing one to sacrifice consistency for computational efficiency.
To understand whether this sacrifice is necessary, recent work developed tools like the feasible subspace dimension to lower bound the prediction dimension of any consistent convex surrogate~\citep{ramaswamy2016convex}.
Challenges of type (II) include estimating risk measures such as conditional value at risk (CVaR), with applications in financial regulation, robust engineering design, and algorithmic fairness.
\jessiet{cites?}
Risk measures are not elicitable, meaning they cannot be specified via a target loss, and thus we seek a surrogate loss of low (or at least finite) prediction dimension.
Recent work~\citep{fissler2016higher,frongillo2015elicitation,frongillo2020elicitation} gives prediction dimension bounds for some of these risk measures, but without the requirement that the surrogate be convex; bounds for convex surrogates are left as a major open question.

@@

\begin{restatable}{theorem}{convexflats}  \label{thm:convex-flats-inf-dim}
  Let $\Gamma:\P \toto \reals^d$ be (directly) elicited by $L \in \Lcvx_d$ for some $d\in\mathbb{N}$.
  Let $\Y$ be either a finite set, or $\Y = \reals$, in which case we assume each $p\in\P$ admits a Lebesgue density supported on the same set for all $p\in\P$.%
  \footnote{This assumption is largely for technical convenience, to ensure that $\V_{u,p}$ does not depend on $p$.
    Any such assumption would suffice, and we suspect even that condition can be relaxed.}
  For all $u\in\range\Gamma$ and $p\in\Gamma_u$, there is some $d$-flat $F$ such that $p \in F \subseteq \Gamma_u$.
 % For all $u\in\range\Gamma$ and $p\in\Gamma_u$, there is some $V_{u,p}:\Y\to\reals^d$ such that $p \in \zeros{V_{u,p}} \subseteq \Gamma_u$.
\raft{Suggestion: let's ditch $V_{u,p}$ in the statement, and instead say ``there is some $d$-flat $F$ such that $p \in F \subseteq \Gamma_u$'', since (a) that's the way we write it everywhere else, (b) we justify the story that our main tool is about $d$-flats, and (c) it becomes clearer why the two Corollaries could follow.  Am I missing a reason to keep $V_{u,p}$? \jessie{I like this change; kept the old version in a comment, but no reason I can think of to keep $V_{u,p}$}}
\end{restatable}
\begin{proof}[Proof (finite case)]
  We will prove the result for the case of finite $\Y$, and defer the $\Y=\reals$ case to \S~\ref{app:implications-of-bounds}.
  As $L$ is convex and elicits $\Gamma$, we have $u \in \Gamma(p) \iff \vec 0 \in \partial \exploss{L}{u}{p}$. 
  With $\Y$ finite, this is additionally equivalent to $\vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the Minkowski sum~\citep[Theorem 4.1.1]{hiriart2012fundamentals}.\footnote{$\partial$ represents the subdifferential $\partial f(x) = \{z : f(x') - f(x) \geq \inprod{z}{x'-x}\; \forall x' \}$.}
  Expanding, we have $\oplus_y p_y \partial L(u,y) = \{ \sum_{y\in\Y} p_y x_y \mid x_y \in \partial L(u,y) \; \forall y\in\Y\}$, and thus there is a $W$ such that $W p = \sum_y p_y x_y = \vec 0$ where $W = [x_1, \ldots, x_n] \in \reals^{d\times n}$; cf.~\cite[$\mathbf{A}^m$ in Theorem 16]{ramaswamy2016convex}.
  Let $V_{u,p} : \Y \to \reals^d, y \mapsto W_y$ be the function encoding the columns of $W$.
  Observe that $\E_p V_{u,p} = \vec 0$.				 
  We take the flat $F := \zeros{V_{u,p}}$, and have $p \in F$ by construction.
  To see $F \subseteq \Gamma_u$, from the chain of equivalences above, we have for any $q\in\P$ that $q \in \zeros{V_{u,p}} \implies \vec 0 \in \partial \E_q L(u,Y) \implies u \in \Gamma(q) \implies q \in \Gamma_u$.
\end{proof}

@@

\paragraph{Q4: Variance}
Consider the task of estimating the variance $\Var(p) = \E_p[Y^2] - \E_p[Y]^2$.
Interestingly, the variance is not elicitable as its level sets are not convex~\cite{osband1985providing,lambert2008eliciting}, meaning this task falls under Quadrant 4.
The fact that the variance is not elicitable does not yield a lower bound on elicitation complexity of 2, as it does not rule out the variance being a link of a real-valued convex-elicitable property; cf.~\citet[Remark 1]{frongillo2020elicitation}.
In \S~\ref{app:variance}, we show $\eliccvx(\Var)=2$, meaning the lowest dimension of a convex loss to estimate conditional variance is 2. %~\citep{osband1985providing,lambert2018elicitation}.
This lower bound will follow from Theorem~\ref{thm:bayes-risk-lower-bound} in \S~\ref{sec:contin-consis} using the fact that variance is the Bayes risk of squared loss. % $L(r,y) = (r-y)^2$, which elicits the mean $\Gamma(p) = \E_p Y$.
Interestingly, while perhaps intuitively obvious, even this simple result is novel.

@@

Using our $d$-flats framework, we resolve both open questions with new lower bounds in both settings.
% The central observation is that the level sets of $d$-dimensional convex-elicitable properties, if not $d$-flats themselves, are
Our framework clarifies the relationship between $d$-identifiable properties and properties elicited by $d$-dimensional convex losses: the level sets of the former are $d$-flats by definition, while the level sets of the latter are \emph{unions} of $d$-flats by Theorem~\ref{thm:convex-flats-inf-dim}.
A careful examination of the arguments of Frongillo and Kash~\citep{frongillo2020elicitation,frongillo2015elicitation} and~\citet{dearborn2020indirect} reveals that they largely rely on the containment of $d$-flats in level sets, rather than the full structure of identifiable properties.
As such, although quite subtle in the case of risk measures, the general structure of these previous proofs go through for convex-elicitable properties: since no $d$-flat could be contained in a particular level set, no union of $d$-flats could be either.
Our lower bounds therefore match both of these papers, though we conjecture that the convex consistency dimension is strictly higher \jessie{than convex identification complexity} in some cases.

@@

The problem of estimating a risk or uncertainty measure of $Y$ is of central importance in financial regulation~\cite{acerbi2014back-testing,fissler2016expected,campbell2005review} and robust engineering design\raft{add cites from AFOSR or EnvAI}.
% , and algorithmic fairness~\citep{williamson2019fairness}
Risk measures include the upper confidence bound $\E[Y] + \lambda \sqrt{\Var[Y]}$, or the Conditional Value at Risk (CVaR) defined below in eq.~\eqref{eq:elic-complex-es-3}, in either conditional or unconditional contexts.
Uncertainty measures include the variance, entropy, or norm of the distribution of $Y$.
Risk and uncertainty measures are typically not elicitable, so this problem falls under Quadrant 4.
Frongillo and Kash~\citep{frongillo2020elicitation,frongillo2015elicitation} give prediction dimension lower bounds for a broad class of risk and uncertainty measures, namely Bayes risks.
As stated above, these bounds are with respect to identifiable properties, and bounds for convex surrogates are left as a major open question.

@@

To illustrate the theorem, we briefly apply it to one of the most prominent financial risk measures, the conditional value at risk (CVaR).
Several other applications from Frongillo and Kash~\citep{frongillo2020elicitation,frongillo2015elicitation}, such as other risk measures, entropy, and norms, follow similarly.
The authors observe that CVaR can be expressed as a Bayes risk; for $0 < \alpha < 1$, we may define
\begin{align}
  \CVaR_\alpha(p)
  &= \inf_{r\in\reals} \E_p\left\{ \tfrac 1 \alpha(r-Y)\ones_{r\geq Y}- r\right\}~,
  \label{eq:elic-complex-es-3}
\end{align}
which is the Bayes risk of the transformed pinball loss $L_\alpha(r,y) = \tfrac 1 \alpha(r-y)\ones_{r\geq y}-r$.
In turn, $L_\alpha$ elicits the $\alpha$-quantile, the quantity $q_\alpha(p)$ such that $\Pr_p[Y \geq q_\alpha(p)] = \alpha$.
Following \citet{frongillo2020elicitation}\jessiet{Leaving as just the 2020 cite}, we will restrict to the set $\P_q$ of probability measures over $\reals$ with connected support and whose CDFs are strictly increasing on their support, so that $q_\alpha$ is single-valued.
Under mild assumptions, we find that there is no consistent real-valued convex surrogate for $\CVaR_\alpha$.

@@

Several important questions remain open.
Particularly for the discrete settings, we would like to know whether one can lift the restriction that surrogates always achieve a minimum; we conjecture positively (see \S~\ref{app:minnable-assumption}).
%For continuous settings, it would be especially interesting to extend to the case $\Y=\reals$, which would require a more careful analysis in Theorem~\ref{thm:bayes-risk-lower-bound}.\jessie{Flagging this bc we take care of it I hope}
Of course, we would like to characterize $\conscvx$ and $\eliccvx$ and develop a general framework for constructing surrogates achieving the best possible bounds for a given notion of prediction dimension.
Moreover, the practical reason why consistency is desired is to ensure the guarantee of empirical risk minimization (ERM) rates; however, the relationship between ERM rates and property elicitation has not been studied.
\jessie{Not sure if there will be a citation for the rates paper...?}
%One might be able to further tighten these bounds by property elicitation by studying monotonicity and adjacency of level sets.
%In discrete predictions, these bounds might also be tightened if the equivalence of convex consistency dimension and embedding dimension of~\citet{finocchiaro2020embedding} is shown; the current embedding dimension bounds are not tight either, but additional structure is imposed by considering the embedding framework.

@@

\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\section{Consistency implies indirect elicitation}\label{sec:consis-implies-indir}

@@

\section{Implications of convex indirect elicitation bounds}\label{app:implications-of-bounds}

@@

\section{Notes on calibration}\label{app:calibration}

When given a discrete target loss, such as for classification-like problems, direct empirical risk minimization is typically NP-hard, forcing one to find a more tractable surrogate.
To ensure consistency, the literature has embraced the notion of \emph{calibration} from~\citet[Chapter 3]{steinwart2008support}, which aligns with the definition in~\citet{tewari2007consistency} for multiclass classification, and its generalizations to arbitrary discrete target losses~\citep{agarwal2015consistent,ramaswamy2016convex}.
Calibration is more tractable and weaker than consistency, yet the two are equivalent under suitable assumptions~\citep{tewari2007consistency,ramaswamy2016convex},notably in Quadrant 1.
Intuitively, calibration says one cannot achieve the optimal surrogate loss while linking to a suboptimal target prediction.

@@

\section{Q1: Previous lower bounds and comparisons}\label{sec:finite-calib}

@@

@@

\section{Omitted discussion and examples}\label{app:omitted}
